---
title: "Developing Data Products 4 Pitch"
author: "Nicholas Vitsentzatos"
date: "4/5/2021"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This presentation is part of the Developing Data Products Coursera.org course project submission.
The Shiny app source code is available at:https://github.com/NicholasVitsentzatos/datasciencecoursera/tree/master/Developing-Data-Products-Assignmet-4

## Main idea of the app

- Your are going to select how many points you want to be displayed on X axis and on the Y axis
- You will have the options whether you want to show the names of the axis and the title of the plot
- After the graph is generated you can brush over the points you want to fit a regression line 

## Code for the R console that will produce similar results

```{r echo = TRUE}
set.seed(123)
number_of_points_we_want=20
dataX <- rnorm(number_of_points_we_want)
dataY <- rnorm(number_of_points_we_want)
df<-data.frame(cbind(dataX,dataY))
fit<-lm(dataY ~ dataX, data =df)
plot(df)
abline(fit, col = "blue", lwd = 2)
```

## Ui / Server Code
```{r,echo=TRUE,eval=FALSE}
ui <- fluidPage(
    titlePanel("Generating random numbers and fitting a regression line on the ones we select"),
    sidebarLayout(
        sidebarPanel(
            numericInput("numeric", "How Many Random Numbers Should be Plotted?", 
                         value = 100, min = 1, max = 200, step = 1),
            sliderInput("sliderX", "Pick Minimum and Maximum X Values",
                        -80, 80, value = c(-50, 50)),
            sliderInput("sliderY", "Pick Minimum and Maximum Y Values",
                        -80, 80, value = c(-50, 50)),
            checkboxInput("show_xlab", "Show/Hide X Axis Label", value = TRUE),
            checkboxInput("show_ylab", "Show/Hide Y Axis Label", value = TRUE),
            checkboxInput("show_title", "Show/Hide Title")
        ),
        mainPanel(
            h3("Please select some points on the graph"),
            plotOutput("plot1",brush = brushOpts(
                id = "brush1"
            ))
        )
    )
)

server <- function(input, output) {
    output$plot1 <- renderPlot({
        set.seed(2016-05-25)
        number_of_points <- input$numeric
        minX <- input$sliderX[1]
        maxX <- input$sliderX[2]
        minY <- input$sliderY[1]
        maxY <- input$sliderY[2]
        dataX <- rnorm(number_of_points, minX, maxX)
        dataY <- rnorm(number_of_points, minY, maxY)
        xlab <- ifelse(input$show_xlab, "X Axis", "")
        ylab <- ifelse(input$show_ylab, "Y Axis", "")
        main <- ifelse(input$show_title, "Graph of the random points :P", "")
        df<-data.frame(cbind(dataX,dataY))
        model <- reactive({
            
            brushed_data <- brushedPoints(df, input$brush1,
                                          xvar = "dataX", yvar = "dataY")
            if(nrow(brushed_data) < 2){
                return(NULL)
            }
            lm(dataY ~ dataX, data = brushed_data)
        })
        plot(df$dataX, df$dataY, xlab = xlab, ylab = ylab, main = main,
             xlim = c(-100, 100), ylim = c(-100, 100))
        if(!is.null(model())){
            abline(model(), col = "blue", lwd = 2)
        }
    })
}
```